---
title: "Modeling in R"
output: html_notebook
---

The Caret Package provides many machine learning models. Install and import:

```{r}
install.packages("caret", dependencies = c("Depends", "Suggests"))
library(caret)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### Load Data

```{r}
data("iris")

dataset <- iris

dataset
```

### Exploratory Data Analysis
```{r}


par(mfrow = c(1,4))
  for(i in 1:4) {
    boxplot(dataset[,i], main=names(iris)[i])
  }
# descriptive statistics
summary(iris)

```

Grouping by Iris type
```{r}
library(ggplot2)
par(mfrow= c(1,4))
  for(i in 1:4){
    boxplot(dataset[,i]~dataset$Species, main=names(iris)[i])
  }

```

Looks like petal length and width are good indicators of the setosa species

Pairplot matrix

```{r}
featurePlot(x=dataset[,1:4], y=dataset[,5],plot="ellipse")
```

### Train Test Split
```{r}
validation_index <- createDataPartition(dataset$Species, p=.75, list=FALSE)
train <- dataset[validation_index,]
test <- dataset[-validation_index,]
```

### Modeling

First set up cross validation and pick metric
```{r}
cv <- trainControl(method = "cv", number = 5)
metric <- "Accuracy"
```

Create models
```{r}
# linear discriminant analysis (lda)
set.seed(123)
fit.lda <- train(Species~., data=train, method= "lda", metric = metric, trControl = cv)

# Classification (and Regression) Trees
set.seed(123)
fit.cart <- train(Species~., data=train, method= "rpart", metric = metric, trControl = cv)

# k Nearest Neighbors
set.seed(123)
fit.knn <- train(Species~., data=train, method= "knn", metric = metric, trControl = cv)

# Support Vector Machine 
set.seed(123)
fit.svm <- train(Species~., data=train, method= "svmRadial", metric = metric, trControl = cv)

#  Random Forrest
set.seed(123)
fit.rf <- train(Species~., data=train, method= "rf", metric = metric, trControl = cv)


# print results
results <- resamples(list(lda = fit.lda, cart = fit.cart, knn = fit.knn, svm = fit.svm, rf = fit.rf))
summary(results)
```
Looks like the linear model did best, but all are highly accurate. Let's check visually
```{r}
dotplot(results)
```

Show best model metrics
```{r}
print(fit.lda)
```
Predict on test set
```{r}
predictions <- predict(fit.lda, test)
confusionMatrix(predictions,test$Species)
```
Model gets 2 incorrect but accuracy doesn't drop too much overall from train to test.